Measurement Bias from Address Aliasing

### 2014-05-02 paper.tex
Small fixes in methodology
Rephrasing and reordering.

% This is basically former background section
Measurement bias has been shown to be commonplace and significant in real applications \cite{Mytkowicz:2008:OE&MB, Mytkowicz:2009:WrongData, Mytkowicz:2008:Easy}.
Previous work looks at environmental factors such as the size of Unix environment variables, which affects memory by offsetting the initial position of stack.
Because of the complexity of modern hardware, it is often impossible to predict exactly how seemingly innocuous things such as environment variables ultimately affect program execution.
While effects of bias has been shown to be commonplace and significant, they are unpredictable and difficult to deal with.
Without an accurate understanding of the underlying memory interactions, the best one can do to accurately assess performance is generally to average over a large sample set of execution contexts.
Previous work focuses mostly of how to account for bias by proper sampling, using randomization of execution contexts and statistical methods. % Cite is basically the same as above, all three are about the same -- difficult to split.

@@

Our setup consists of an Intel\textregistered{} Core\texttrademark{} i7-4770K CPU @ 3.50 GHz, running 64-bit Ubuntu 13.10.
Because bias can be hardware dependent, and to keep the scope of this work manageable, we choose to focus only on this processor and the Intel ``Haswell'' microarchitecture specifically.
Code samples are compiled using the GCC toolchain, version 4.8.1-10ubuntu9.

We also have to ensure we are not affected by measurement bias beyond what we are actually trying to observe, and in general follow best practices on gathering data \cite{Mytkowicz:2009:WrongData}.
Most importantly, this means keeping the memory address space under control.
For security reasons, addresses of stack, heap and dynamic libraries are often randomized at load time, a technique known as Address Space Layout Randomization.\cite{Shackham:2004:ASLR} % Should maybe have more generic reference
By disabling \emph{ASLR}, we are able to execute the same program multiple times with identical virtual address spaces.
All benchmarks are done on a machine under minimal load, and \emph{Frequency scaling} is disabled to keep the processor's clock speed fixed.
Finally, we disable \emph{Hyper-threading}, which reduces the possibility for resource contention between threads.

>>>>

% This is basically former background section
Measurement bias has been shown to be commonplace and significant in real applications~\cite{Mytkowicz:2008:OE&MB, Mytkowicz:2009:WrongData, Mytkowicz:2008:Easy}.
Previous work looks at environmental factors such as the size of Unix environment variables, which affects memory by offsetting the initial position of stack.
Because of the complexity of modern hardware, it is often impossible to predict exactly how seemingly innocuous things such as environment variables ultimately affect program execution.
While effects of bias has been shown to be commonplace and significant, they are unpredictable and difficult to deal with.
Without an accurate understanding of the underlying memory interactions, the best one can do to accurately assess performance is generally to average over a large sample set of execution contexts.
Previous work focuses mostly of how to account for bias by proper sampling, using randomization of execution contexts and statistical methods. % Cite is basically the same as above, all three are about the same -- difficult to split.

@@

Because bias can be hardware dependent, and to keep the scope of this work manageable, we choose to focus on the Intel ``Haswell'' microarchitecture specifically.
Our setup consists of a 4th generation Intel\textsuperscript{\textregistered} Core\textsuperscript{\texttrademark} i7-4770K processor, running 64-bit Ubuntu 13.10.
Code samples are compiled using the GCC toolchain, version 4.8.1-10ubuntu9.

We also have to ensure we are not affected by measurement bias beyond what we are actually trying to observe, and in general follow best practices on gathering data~\cite{Mytkowicz:2009:WrongData}.
Most importantly, this means keeping the memory address space under control.
For security reasons, addresses of stack, heap and dynamic libraries are often randomized at load time, a technique known as \emph{Address Space Layout Randomization}~\cite{Shackham:2004:ASLR}. % Should maybe have more generic reference
By disabling ASLR, we are able to execute the same program multiple times with identical virtual address spaces.
All benchmarks are done on a machine under minimal load, and \emph{frequency scaling} is disabled to keep the CPU's clock speed fixed.
Finally, we disable \emph{Hyper-threading}, which reduces the possibility for resource contention between threads.
